---
title: "FAO.FishStat.Global Capture Production.Data"
author: "Thomas Berger"
date: "1/30/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## FAO FishStat Global Catch 2014

FAO Global catch contains capture production statistics by country or territory, species item, and FAO Major Fishing Area for the reporting year 2014; the most recent year for which data is available at the time of writing this report.

The github repository is here:
https://github.com/bergertom/Plotly

The plotly slide show can be accessed here:
https://bergertom.github.io/Plotly/

## Data source

The data was loaded from the FAO query panel: http://www.fao.org/fishery/statistics/global-capture-production/query/en, and saved into a CSV file for this report. Data downloaded 09/01/2017.

Data shown is the global catch for human consumption; reported by FAO under as [Fish, crustaceans and molluscs, etc.] group.

To arrive at the detail, select country [Alphabetical], on the Index select [All] to have all countries, then click [select all] to include them all into the query. Under time, select [2014] as year. Under Species, select [Ybk main groups], select [Fish, crustaceans and molluscs, etc.] to have data for human consumption. Under Display, select [Country] and [Quantity] to have the result by country.


## Read and clean the data

```{r echo=TRUE}
   # read data downloaded fom FAO web site
   faodata <- read.csv("figis_guestnull.csv", na.strings = c("NA", "", NULL))
   # remove numbers
   faodata <- subset(faodata, Unit == 't')
   # remove unit
   faodata$Unit <- NULL
   # remove symbol
   faodata$S_2014 <- NULL
   # fix column names
   colnames(faodata)<- c("Country","Value.2014")
   str(faodata)
```

## Prepare data for the plot
```{r echo=TRUE}
   # rank all countries catch
   faodata <- transform(faodata, 
          rank = ave(Value.2014, 
                          FUN = function(x) rank(-x, ties.method = "first")))
   # add ranking 11 ... 233 together
   other_value <- sum(faodata[faodata$rank > 10,"Value.2014"])
   # leave only the top 10 ranking countries
   faodata <- subset(faodata, rank <= 10)
   # add the others
   faodata$rank <- NULL
   faodata <- rbind(faodata, data.frame(Country="Others", Value.2014=other_value))
   # shorten names for plot
   faodata$Country <- as.character(faodata$Country)
   faodata[9,1] <- "USA"
   faodata[8,1] <- "Russia"
   # round to millions
   faodata$Value.2014 <- round(faodata$Value.2014 / 1000000,2)
```

## Global Capture 2014

```{r warning=FALSE, echo=FALSE, message=FALSE}
   library(plotly)
   p <- plot_ly(faodata, x = ~Country, y = ~Value.2014, name = "Women",
             mode = "markers", marker = list(color = "blue")) %>%
      layout(
        title = "top 10 countries",
        yaxis = list(title = "Catch (in millions) tons"),
        margin = list(l = 65)
   )
   p
```
[FAO.FishStat.Global Capture Production.Data: 2014](http://www.fao.org/fishery/statistics/global-capture-production/query/en).

