---
title: "FAO.FishStat.Global Capture Production.Data"
author: "Thomas Berger"
date: "1/16/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## FAO Global Catch 2014

FAO Global catch contains capture production statistics by country or territory, species item, and FAO Major Fishing Area for the reporting year 2014.

The data was loaded from the FAO query panel: http://www.fao.org/fishery/statistics/global-capture-production/query/en, and saved into a CSV file for this report. Data downloaded 09/01/2017.

The github repository is here:
https://github.com/bergertom/Plotly

The plotly slide show can be accessed here:
https://bergertom.github.io/Plotly/

## Read and clean the data

```{r echo=TRUE}
   # read data downloaded fom FAO web site
   faodata <- read.csv("figis_guestnull.csv", na.strings = c("NA", "", NULL))
   # remove numbers
   faodata <- subset(faodata, Unit == 't')
   # remove unit
   faodata$Unit <- NULL
   # remove symbol
   faodata$S_2014 <- NULL
   # fix column names
   colnames(faodata)<- c("Country","Value.2014")
   str(faodata)
```

## Prepare data for the plot
```{r echo=TRUE}
   # rank all countries catch
   faodata <- transform(faodata, 
          rank = ave(Value.2014, 
                          FUN = function(x) rank(-x, ties.method = "first")))
   # add ranking 11 ... 233 together
   other_value <- sum(faodata[faodata$rank > 10,"Value.2014"])
   # leave only the top 10 ranking countries
   faodata <- subset(faodata, rank <= 10)
   # add the others
   faodata$rank <- NULL
   faodata <- rbind(faodata, data.frame(Country="Others", Value.2014=other_value))
   # shorten names for plot
   faodata$Country <- as.character(faodata$Country)
   faodata[9,1] <- "USA"
   faodata[8,1] <- "Russia"
   # round to millions
   faodata$Value.2014 <- round(faodata$Value.2014 / 1000000,0)
```

## Global Capture 2014

```{r warning=FALSE}
   library(ggplot2)
   ggplot(faodata, aes(x=Country, y=Value.2014)) +
       geom_point() +
       ggtitle("Fishery: Global Catch 2014") +
       labs(x="Country",y="Tons (million)") 
```
[FAO.FishStat.Global Capture Production.Data: 2014](http://www.fao.org/fishery/statistics/global-capture-production/query/en).

